Improving Software Maintainability through Automated Refactoring of Code Clones

### 2019-08-24 icse/paper.tex
Replaced all results
Aaah I don't know what to do with this thresholds thing

\subsubsection{Extract Method}
% Show my categories to show what clones can be dealt with by method extraction.
Several influencing factors may obstruct the possibility to extract code to a new method:
\begin{itemize}
    \item \textbf{Complex Control Flow}: This clone contains \texttt{break}, \texttt{continue} or \texttt{return} statements, obstructing the possibility of method extraction.
    \item \textbf{Spans Part Of A Block}: This clone spans a part of a block statement.
    \item \textbf{Is Not A Partial Method}: If the clone does not fall in the ``Partial method'' category of Sec.~\ref{sec:location}, the ``Extract Method'' refactoring technique cannot be applied.
    \item \textbf{Returns Multiple Values}: If a clone modifies or declares multiple pieces of data that it should return.
    \item \textbf{Top-Level Non-Statement}: If one of the top-level AST nodes of the clone is not a statement. For instance, if a (part of) an anonymous class is cloned.
    \item \textbf{Can Be Extracted}: This clone is a fragment of code that can directly be extracted to a new method. Then, based on the relation between the clone instances, further refactoring techniques can be used to refactor the extracted methods (for instance ``pull up method'' for clones in sibling classes).
\end{itemize}
Clones that do not fall in the \textit{Can Be Extracted} category may require additional transformations or other techniques to refactor. CloneRefactor only refactors the clones that \textit{Can Be Extracted}.

@@

\subsection{Minimum clone size}
In this study, we want to find out what thresholds to use to improve maintainability if clones by those thresholds are refactored. However, when clones are very small, they may never be able to improve maintainability. The detrimental effect of the added volume of the newly created method exceeds the positive effect of removing duplication. Because of that, we perform all our experiments with a minimum clone size of 10 tokens, because smaller clones cannot improve maintainability when refactored.

@@

@@

\subsection{Extract Method}
%To what extent can found clones be refactored through method extraction, without requiring additional transformations.
Table~\ref{tab:refactorability} shows to what extent clone classes can be refactored by using the ``Extract Method'' refactoring technique (see ). %The second column shows our measurements for the complete systems (just like the former experiments). %The third column shows our measurements when restricting our search to method bodies. The amount that can be extracted increases because, when restricting our search to method bodies, we do not exclude declarations that can obstruct the possibility of method extraction (for instance a cloned method signature).

@@

\subsection{Refactoring}
%I think the ultimate goal with this thesis is to do experiments with different clone thresholds. Which thresholds give clones that we should refactor? For this, we will measure the maintainability of the refactored source code over different thresholds. These thresholds range from minimum clone size, variability, and gap size.
In our corpus, CloneRefactor has refactored 12.710 clone classes and measured the change in indicated metrics (see Sec.~\ref{sec:metrics}). Using the presented formulas (see Sec.~\ref{sec:metricformula}) we determine how the characteristics of the extracted method (see Sec.~\ref{sec:characteristics}) influence the maintainability of the resulting codebase after refactoring. In this section, we explore the data received by comparing the before- and after snapshots of the system for each separate refactoring.

@@

\begin{table*}[htb]
\centering
%\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrrrrr@{}}
\toprule
\textit{\textbf{Relation}} & \textit{\textbf{Duplication}} & \textit{\textbf{Complexity}} & \textit{\textbf{Parameters}} & \textit{\textbf{Volume}} & \textit{\textbf{\#}} & \textit{\textbf{Score}} \\ \midrule
\textbf{Common Hierarchy} & \textbf{-66.33} & \textbf{0.73} & \textbf{1.20} & \textbf{-8.85} & \textbf{2,202} & \textbf{0.23} \\ \midrule
\hspace{10pt} Superclass & -64.48 & 0.79 & 0.94 & -7.22 & 229 & 0.42 \\
\hspace{10pt} Sibling & -70.07 & 0.69 & 1.28 & -10.97 & 1,722 & 0.23 \\
\rowcolor[HTML]{FFCCC9}
\hspace{10pt} Same Hierarchy & -44.18 & 0.95 & 0.89 & 1.54 & 87 & 0.10 \\
\hspace{10pt} First Cousin & -42.69 & 0.89 & 0.93 & 4.86 & 144 & 0.02 \\
\rowcolor[HTML]{FFCCC9}
\hspace{10pt} Ancestor & -32.75 & 1.00 & 0.75 & 11.00 & 20 & -0.03 \\ \midrule
\textbf{Common Interface} & \textbf{-47.06} & \textbf{0.83} & \textbf{1.04} & \textbf{4.50} & \textbf{1,044} & \textbf{-0.02} \\ \midrule
\hspace{10pt} Same Indirect Interface & -37.08 & 0.93 & 0.82 & 9.96 & 487 & -0.01 \\
\hspace{10pt} Same Direct Interface & -55.79 & 0.75 & 1.24 & -0.28 & 557 & -0.02 \\ \midrule
\textbf{Common Class} & \textbf{-52.42} & \textbf{0.87} & \textbf{1.13} & \textbf{1.47} & \textbf{7,239} & \textbf{-0.02} \\ \midrule
\hspace{10pt} Same Class & -51.85 & 0.86 & 1.03 & 3.36 & 4,874 & 0.04 \\
\hspace{10pt} Same Method & -53.60 & 0.90 & 1.32 & -2.44 & 2,365 & -0.15 \\ \midrule
\textbf{Unrelated} & \textbf{-45.86} & \textbf{0.88} & \textbf{1.08} & \textbf{9.56} & \textbf{2,198} & \textbf{-0.15} \\ \midrule
\hspace{10pt} No Direct Superclass & -52.24 & 0.84 & 1.12 & 6.04 & 811 & -0.06 \\
\hspace{10pt} External Superclass & -47.09 & 0.87 & 1.13 & 8.77 & 697 & -0.17 \\
\hspace{10pt} External Ancestor & -35.73 & 0.93 & 0.95 & 14.58 & 586 & -0.21 \\
\hspace{10pt} No Indirect Superclass & -44.89 & 0.84 & 1.18 & 14.08 & 104 & -0.30 \\ \midrule
\textbf{Grand Total} & \textbf{-53.26} & \textbf{0.84} & \textbf{1.12} & \textbf{1.33} & \textbf{12,683} & \textbf{0.00} \\ \bottomrule
\end{tabular}%
%}
\caption{Influence on maintainability of refactoring clones with the specified relation categories.}
\label{tab:relation_refactor}
\end{table*}

>>>>

\subsubsection{Extract Method} \label{sec:refactorability}
% Show my categories to show what clones can be dealt with by method extraction.
Several influencing factors may obstruct the possibility to extract code to a new method:
\begin{itemize}
    \item \textbf{Complex Control Flow}: This clone contains \texttt{break}, \texttt{continue} or \texttt{return} statements, obstructing the possibility of method extraction.
    \item \textbf{Spans Part Of A Block}: This clone spans a part of a block statement.
    \item \textbf{Is Not A Partial Method}: If the clone does not fall in the ``Partial method'' category of Sec.~\ref{sec:location}, the ``Extract Method'' refactoring technique cannot be applied.
    \item \textbf{Returns Multiple Values}: If a clone modifies or declares multiple pieces of data that it should return.
    \item \textbf{Top-Level Non-Statement}: If one of the top-level AST nodes of the clone is not a statement. For instance, if a (part of) an anonymous class is cloned.
    \item \textbf{Can Be Extracted}: This clone is a fragment of code that can directly be extracted to a new method. Then, based on the relation between the clone instances, further refactoring techniques can be used to refactor the extracted methods (for instance ``pull up method'' for clones in sibling classes).
\end{itemize}
Clones that do not fall in the \textit{Can Be Extracted} category may require additional transformations or other techniques to refactor. CloneRefactor only refactors the clones that \textit{Can Be Extracted}.

@@

\subsection{Minimum clone size}
In this study, we want to find out what thresholds to use to improve maintainability if clones by those thresholds are refactored. However, when clones are very small, they may never be able to improve maintainability. The detrimental effect of the added volume of the newly created method exceeds the positive effect of removing duplication. Because of that, we perform all our experiments with a minimum clone size of 10 tokens, because smaller clones are very unlikely to improve maintainability when refactored.

\subsection{Thresholds}
Most clone detection tools can be configured using thresholds. These thresholds indicate the minimum number of lines, tokens and/or statements that must be spanned for duplicate fragments to be considered clones. Often, such thresholds are intuitively chosen \cite{li2006cp, roy2009mutation} or based on a quartile distribution of empirical data \cite{alves2010deriving}. Using the maintainability score we can find support for which thresholds should be chosen to increase the chance to find clones that improve maintainability when refactored.

%In our experiments, we compare the maintainability scores for different thresholds. We use the \textit{standard error} to determine the correlation between the used threshold metric and the maintainability score. A low standard error shows that the threshold metric has a high influence on the maintainability score. This makes it more suitable to use as a threshold. By plotting the maintainability scores for different thresholds, we can determine which thresholds should be used to positively influence maintainability.

%We define the following volume-based thresholds for clones, ordered by how often clone detection tools support such thresholds:
%\begin{itemize}
%\item \textbf{Number of lines}: The number of lines a single instance of the clone class spans.
%\item \textbf{Tokens}: The number of tokens a single instance of the clone class spans.
%\item \textbf{Number of statements}: The number of statements a single clone instance. Often closely reflects the number of lines, but is not dependent on coding style (number of newlines used by the programmer in statements).
%\item \textbf{Token Volume}: The number of tokens in all clone instances in a clone class combined.
%\end{itemize}

@@

%% RESULTS TABLES (MOVED UP SO LATEX PLACES THEM IN THE RIGHT PLACE)

\begin{table*}
\centering
%\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrrrrr@{}}
\toprule
\textit{\textbf{Relation}} & \textit{\textbf{Duplication}} & \textit{\textbf{Complexity}} & \textit{\textbf{Parameters}} & \textit{\textbf{Volume}} & \textit{\textbf{\#}} & \textit{\textbf{Score}} \\ \midrule
\textbf{Common Hierarchy} & \textbf{-66.33} & \textbf{0.73} & \textbf{1.20} & \textbf{-8.85} & \textbf{2,202} & \textbf{0.23} \\ \midrule
\hspace{10pt} Superclass & -64.48 & 0.79 & 0.94 & -7.22 & 229 & 0.42 \\
\hspace{10pt} Sibling & -70.07 & 0.69 & 1.28 & -10.97 & 1,722 & 0.23 \\
\rowcolor[HTML]{FFCCC9}
\hspace{10pt} Same Hierarchy & -44.18 & 0.95 & 0.89 & 1.54 & 87 & 0.10 \\
\hspace{10pt} First Cousin & -42.69 & 0.89 & 0.93 & 4.86 & 144 & 0.02 \\
\rowcolor[HTML]{FFCCC9}
\hspace{10pt} Ancestor & -32.75 & 1.00 & 0.75 & 11.00 & 20 & -0.03 \\ \midrule
\textbf{Common Interface} & \textbf{-47.06} & \textbf{0.83} & \textbf{1.04} & \textbf{4.50} & \textbf{1,044} & \textbf{-0.02} \\ \midrule
\hspace{10pt} Same Indirect Interface & -37.08 & 0.93 & 0.82 & 9.96 & 487 & -0.01 \\
\hspace{10pt} Same Direct Interface & -55.79 & 0.75 & 1.24 & -0.28 & 557 & -0.02 \\ \midrule
\textbf{Common Class} & \textbf{-52.42} & \textbf{0.87} & \textbf{1.13} & \textbf{1.47} & \textbf{7,239} & \textbf{-0.02} \\ \midrule
\hspace{10pt} Same Class & -51.85 & 0.86 & 1.03 & 3.36 & 4,874 & 0.04 \\
\hspace{10pt} Same Method & -53.60 & 0.90 & 1.32 & -2.44 & 2,365 & -0.15 \\ \midrule
\textbf{Unrelated} & \textbf{-45.86} & \textbf{0.88} & \textbf{1.08} & \textbf{9.56} & \textbf{2,198} & \textbf{-0.15} \\ \midrule
\hspace{10pt} No Direct Superclass & -52.24 & 0.84 & 1.12 & 6.04 & 811 & -0.06 \\
\hspace{10pt} External Superclass & -47.09 & 0.87 & 1.13 & 8.77 & 697 & -0.17 \\
\hspace{10pt} External Ancestor & -35.73 & 0.93 & 0.95 & 14.58 & 586 & -0.21 \\
\hspace{10pt} No Indirect Superclass & -44.89 & 0.84 & 1.18 & 14.08 & 104 & -0.30 \\ \midrule
\textbf{Grand Total} & \textbf{-53.26} & \textbf{0.84} & \textbf{1.12} & \textbf{1.33} & \textbf{12,683} & \textbf{0.00} \\ \bottomrule
\end{tabular}%
%}
\caption{Influence on maintainability of refactoring clones with the specified relation categories.}
\label{tab:relation_refactor}
\end{table*}

%% END TABLES

@@

\subsection{Refactorability}
%To what extent can found clones be refactored through method extraction, without requiring additional transformations.
Table~\ref{tab:refactorability} shows to what extent clone classes can be refactored by using the ``Extract Method'' refactoring technique (see Sec.~\ref{sec:refactorability}). %The second column shows our measurements for the complete systems (just like the former experiments). %The third column shows our measurements when restricting our search to method bodies. The amount that can be extracted increases because, when restricting our search to method bodies, we do not exclude declarations that can obstruct the possibility of method extraction (for instance a cloned method signature).

@@

These results indicate that, given our refactorability criteria, 28\% of clones can be automatically refactored. Clones in other categories may require other refactoring techniques or further transformations to be automatically refactorable.

\subsection{Thresholds}
%I think the ultimate goal with this thesis is to do experiments with different clone thresholds. Which thresholds give clones that we should refactor? For this, we will measure the maintainability of the refactored source code over different thresholds. These thresholds range from minimum clone size, variability, and gap size.
In our corpus, CloneRefactor has refactored 12.710 clone classes and measured the change in indicated metrics (see Sec.~\ref{sec:metrics}). Using the presented formulas (see Sec.~\ref{sec:metricformula}) we determine how the characteristics of the extracted method (see Sec.~\ref{sec:characteristics}) influence the maintainability of the resulting codebase after refactoring. In this section, we explore the data received by comparing the before- and after snapshots of the system for each separate refactoring. Using this data, we find supporting evidence regarding which thresholds are most likely to find clones that should be refactored to improve maintainability.

@@
